---
title: "Course Project One"
author: "Justin Bartee"
date: "June 19, 2018"
output: html_document
---

```{r setup, include=FALSE}
library(ggplot2)
knitr::opts_chunk$set(echo = TRUE)
```

## Summary

This document is composed of the analysis and figures meeting the requirements for submission for the Course Project One assignment.

## Data
Data from the accompanying "activity.zip" file are loaded:

```{r}
activitydata <- read.csv(unz("activity.zip", "activity.csv"))
activitydata$date <- as.Date(activitydata$date)
```

## Analysis of Steps Per Day

We examine the data aggregated by day, specifically mean and median values, ignoring missing values:
```{r}
dailysteps <- aggregate(steps ~ date, data=activitydata, FUN=sum, na.action = na.omit)
meandaily <- mean(dailysteps$steps)
mediandaily <- median(dailysteps$steps)
```

The mean is `r format(meandaily[1],scientific=FALSE)`.
The median is `r mediandaily[1]`.


A histogram of the daily step count provides some insight into this:
```{r}
dailyhist <- qplot(dailysteps$steps, 
                   geom="histogram",
                   fill=I("lightblue"),
                   col=I("red"),
                   binwidth=5000,
                   main = "Daily Steps Histogram",
                   xlab = "Steps",
                   ylab = "Days") +
  geom_vline(xintercept=meandaily,color="yellow") +
  geom_vline(xintercept=mediandaily,color="black")
dailyhist
```

