---
title: "Course Project One"
author: "Justin Bartee"
date: "June 19, 2018"
output: html_document
---

```{r setup, include=FALSE}
library(ggplot2)
knitr::opts_chunk$set(echo = TRUE)
```

## Summary

This document is composed of the analysis and figures meeting the requirements for submission for the Course Project One assignment.

## Data
Data from the accompanying "activity.zip" file are loaded:

```{r}
activitydata <- read.csv(unz("activity.zip", "activity.csv"))
activitydata$date <- as.Date(activitydata$date)
```

## Analysis of Steps Per Day

We examine the data aggregated by day, specifically mean and median values, ignoring missing values:
```{r}
dailysteps <- aggregate(steps ~ date, data=activitydata, FUN=sum, na.action = na.omit)
meandaily <- mean(dailysteps$steps)
mediandaily <- median(dailysteps$steps)
```

The mean is `r format(meandaily[1],scientific=FALSE)`.

The median is `r mediandaily[1]`.


A histogram of the daily step count may provide additional insight into this:
```{r}
dailyhist <- qplot(dailysteps$steps, 
                   geom="histogram",
                   fill=I("lightblue"),
                   col=I("red"),
                   binwidth=5000,
                   main = "Daily Steps Histogram",
                   xlab = "Steps",
                   ylab = "Days") +
  geom_vline(xintercept=meandaily,color="yellow") +
  geom_vline(xintercept=mediandaily,color="black")
dailyhist
```

The median and mean are quite close, almost overlapping, as shown on the plot by the vertical lines.


## Time Series

We might gain additional insight by examining the time series for various intervals. For this project, the 5-minute interval is chosen.

We acquire the intervals by aggregation, ignoring missing values:
```{r}
intervalsteps <- aggregate(steps ~ interval, data=activitydata, FUN=mean)
```

And examine a plot of the time series:

```{r}
intervalplot <- ggplot(intervalsteps, aes(interval, steps)) +
  geom_line() +
  xlab("Interval") +
  ylab("Steps (mean per day)")
intervalplot
```

We compute some information related to the interval with the maximum value:
```{r}
maxindex <- which(intervalsteps$steps == max(intervalsteps$steps))
intervalmax <- intervalsteps$interval[maxindex]
maxsteps <- intervalsteps$steps[maxindex]
intervalhour <- floor(intervalmax / 60) + 1
```
We see that the interval with the maximum value is `r intervalmax` (hour `r intervalhour`). The maximum average steps taken during this interval is about `r floor(maxsteps)`. 

## Handling Missing Data
